Introduction to the Tidyverse

```{r}
install.packages("gapminder")
install.packages("dplyr")

library(gapminder)
library(dplyr)

gapminder

```
Filtering for one year

```{r}
gapminder %>%
  filter(year==1957)
```
Filtering for one country and one year

```{r}
gapminder %>%
  filter(year==1957, country=="China")
```

Arranging observations by life expectancy

```{r}
gapminder %>%
  arrange(lifeExp)

# sort in descending order of lifeExp

gapminder %>%
  arrange(desc(lifeExp))
```
Filtering and arranging
```{r}
gapminder %>%
  filter(year==1957) %>%
  arrange(desc(pop))
```
Using mutte to change or create a column

```{r}
#Use mutate to change lifeExp to be in months

gapminder %>%
  mutate(lifeExp = 12*lifeExp)

#Use mutate to create a new column called lifeExpMonths

gapminder %>%
  mutate(lifeExpMonths = 12*lifeExp)

```
Combining filter, mutate and arrange

```{r}
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12*lifeExp) %>%
  arrange(desc(lifeExpMonths))
```
Visualizing with ggplot2
```{r}
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) + 
  geom_point()

```
```{r}
library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) + 
  geom_point()

```
Putting the x-axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) +
  geom_point() +
  scale_x_log10()
```
Putting the x-axes and y-axes on a log scale

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()
```
Adding color to a scatter plot

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent)) +
  geom_point() + 
  scale_x_log10()
```

Adding size and color to a plot
```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent, size=gdpPercap)) +
  geom_point() + 
  scale_x_log10()
```

Faceting

Creating a subgragh for each continent

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) +
  geom_point() + 
  scale_x_log10() +
  facet_wrap(~ continent)
```

Faceting by year

```{r}
ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color=continent, size= pop)) +
  geom_point() + 
  scale_x_log10() +
  facet_wrap(~ year)
```

