Grouping and summarizing

Summarizing the median life expectancy

```{r}
gapminder %>%
  summarize(medianLifeExp = median(lifeExp))

#Filter for 1957 then summarize the median life expectancy

gapminder %>%
  filter(year==1957) %>%
  summarize(medianLifeExp = median(lifeExp))

gapminder %>%
  filter(year==1957) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))


```

Summarizing by year

```{r}
gapminder %>%
  group_by(year) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```
Summarizing by continent

```{r}
gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap)) %>%
  arrange(desc(maxGdpPercap))
```

Summarizing by continent and year

```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```

Line Plots

Visualizing median life expectancy over time

```{r}
by_year <- gapminder %>%
  group_by(year) %>%
  summarise(medianGdpPercap = median(gdpPercap))

by_year
ggplot(by_year, aes(x=year, y=medianGdpPercap)) +
  geom_line()
```
```{r}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_year_continent

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color=continent)) +
  geom_line()
```
Bar Plots

Visualizing median GDP per capita by continent
```{r}

by_continent <- gapminder %>%
  filter(year==1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))
            
by_continent

ggplot(by_continent, aes(x=continent, y=medianGdpPercap)) +
  geom_col()

```
Visualizing GDP per capita by countru in Oceania

```{r}
oceania_1952 <- gapminder %>%
  filter(continent=="Oceania", year==1952)

oceania_1952

ggplot(oceania_1952, aes(x=country, y= gdpPercap))+
  geom_col()
```


Histogram

Visualizing population

```{r}
library(gapminder)
library(dplyr)
library(ggplot2)

gapminder

gapminder_1952 <- gapminder %>%
  filter(year==1952) %>%
  mutate(pop_by_mil = pop/1000000)

gapminder_1952

ggplot(gapminder_1952, aes(x=pop_by_mil)) +
  geom_histogram(bins = 50)
```

Visualizing population with x-axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop)) + 
  geom_histogram(bins = 50) +
  scale_x_log10()

```

Boxplots

Comparing GDP per capita  across continents

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

gapminder_1952 %>%
  filter(continent=="Africa")

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() + 
  scale_y_log10()
```
Adding a title to your graph
```{r}
ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() + 
  scale_y_log10() +
  ggtitle("Comparing DPG per capita across continents")
```

Data Manipulation with dplyr

```{r}
library(tidyverse)
counties <- read_csv("counties.csv")
glimpse(counties)
counties
```

Selecting columns

```{r}
counties %>%
  select(state, county, population,  men, women)

counties_Selected <-counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_Selected %>%
  arrange(desc(public_work))

counties_Selected <- counties %>%
  select(state, county, population)

counties_Selected %>%
  filter(population >1000000)

counties_Selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_Selected %>%
  filter(state=="Texas", population >10000) %>%
  arrange(desc(private_work))
```
Select, mutate, filter and arrange

```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population > 10000) %>%
  arrange(desc(proportion_men))

```
 The verb count
 
```{r}
select_counties <- counties %>%
  select(state, region, county, population, citizens)

counties %>%
  count(region, sort=TRUE) %>%
  arrange(desc(n))
```
 Counting citizens by state
 
```{r}
counties_selected <- counties %>%
  select(county, region, state, population, citizens)

counties_selected %>%
  count(state, wt=citizens) %>%
  arrange(desc(n))
```
 
Mutating and counting

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, walk)

counties_selected %>%
  mutate(population_walk = population*walk/100) %>%
  count(state, wt=population_walk, sort=TRUE)

```
Summarizing

```{r}
counties_selected <- counties %>%
  select(county, population, income, unemployment)

counties_selected %>%
  summarize(min_population = min(population), max_unemployment=max(unemployment), averagen_income=mean(income))
```

Summarizing by state

```{r}
counties_selected <- counties %>%
  select(state, county, population, land_area)

counties_selected %>%
  group_by(state) %>%
  summarize(total_area=sum(land_area), total_population=sum(population)) %>%
  mutate(density=total_population/total_area) %>%
  arrange(desc(density))

counties_selected <- counties %>%
  select(region, state, county, population)

counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) %>%
  summarize(average_pop=mean(total_pop), median_pop=median(total_pop))
```
Top_n verb

```{r}
counties_selected <- counties %>%
  select(region, state, county, metro, population, walk)

counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)

counties_selected <- counties %>%
  select(region, state, county, population, income)

counties_selected %>%
  group_by(region, state) %>%
  summarize(average_income=mean(income)) %>%
  top_n(1, average_income)
```

```{r}
counties_selected <- counties %>%
  select(state, metro, population)

counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop=sum(population)) %>%
  top_n(1,total_pop) %>%
  ungroup() %>%
  count(metro)
```
Selecting columns

```{r}
counties %>%
  select(state, county, population, professional:production) %>%
  arrange(desc(service))

counties %>%
  select(state, county, population, ends_with("work")) %>%
  filter(public_work >=50)
```

